---
title: "Introduction to Deduplication and Weighted Deduplication by RecordLinkage"
author: "Hathaway Liu, Qinzhi Peng"
date: "2024-10-01"
output: html_document
---

```{r}
library(RecordLinkage)
```

```{r}
data(RLdata500)
RLdata500[1:5, ]
```

### Comparison Pattern

RecordLinkage contains two functions that could create comparison pattern of dataset:

1.  compare.dedup: To compare records within a single dataset to identify duplicates

    `compare.dedup (dataset, blockfld = FALSE, phonetic = FALSE,`

    `phonfun = soundex, strcmp = FALSE, strcmpfun = jarowinkler, exclude = FALSE,`

    `identity = NA, n_match = NA, n_non_match = NA)`

2.  compare.linkage: To compare records across two different datasets to find matches

    `compare.linkage (dataset1, dataset2, blockfld = FALSE,`

    `phonetic = FALSE, phonfun = soundex, strcmp = FALSE,`

    `strcmpfun = jarowinkler, exclude = FALSE, identity1 = NA, identity2 = NA,`

    `n_match = NA, n_non_match = NA)`

The key difference between these two functions is finding matches within a single dataset or across two different datasets.

```{r}
rpairs <- compare.dedup(RLdata500, identity = identity.RLdata500)
rpairs$pairs[1:5, ]
```

By calling `rpairs$pairs` , we could extract the comparison pattern resulted from `compare.dedup`

In above example, NA usually indicates the missing value in the dataset. id1 and id2 are two person that being compared; wherever there's a match, there will be a 1 indicating a match for specific column. For instance, person with id = 1 and person with id = 2 have a match on birth month `bm` . The final column `is_match` indicates if two persons are the same (1 for match, 0 for non-match)

There are more parameters we could tune to use above functions in different cases:

-   `blockfld`: Specifies the fields on which blocking should occur to limit the number of record comparisons (e.g., blocking on names)

-   `strcmp`: determines whether or not string comparison should be performed (such as `strcmp = TRUE` for names).

-   `phonetic`, `phonfun`: determines if a phonetic algorithm is used to compare strings (e.g. names) that sound alike but different in spelling; when `phonetic = TRUE`, phonetic algorithm applies to all columns, or giving specific column number to apply for chosen column. Specify phonetic algorithm to use using `phonfun`

-   `exclude`: specify if certain fields should be excluded from comparison

-   `identity`: store the true identity of records for validation

-   `n_match`, `n_non_match`: set manually the number of record pairs that are identified as matches or non-matches

More detailed usage of different parameters will be discussed in the later sections:

### Blocking

Blocking is a technique used in record linkage and deduplication to reduce the number of comparisons made between records. Every pair of records will be compared without blocking, which result in a long and expensive computation.

Thus, blocking helps by narrowing down the number of comparisons, where it constructs **blocking fields** that only compares records that share the same values for certain fields. There could be a single blocking field, or combining multiple fields, but only records that match in all specified fields will be compared.

Blocking could be easily implemented by utilizing `blockfld` parameter mentioned above:

1.  Single blocking field: only records with the same first name will be compared

```{r}
rpairs <- compare.dedup(RLdata500, blockfld = 1, identity = identity.RLdata500)
rpairs$pairs[1:5, ]
```

2.  Multiple blocking fields: Records with same birth dates will be compared.

```{r}
rpairs <- compare.dedup(RLdata500, blockfld = c(5:7), identity = identity.RLdata500)
rpairs$pairs[1:5, ]
```

3.  Combination: Records with same last name and same birth day and month will be compared.

```{r}
rpairs <- compare.dedup(RLdata500, blockfld = list(3, 6:7), identity = identity.RLdata500)
rpairs$pairs[c(1:9, 10:11), ]
```

### Phonetic Functions and String Comparison

Phonetic functions and string comparators are two methods used in record linkage to handle variations or errors in text data, like spelling mistakes or differences in pronunciation.

**Phonetic functions** help match words that sound the same but might be spelled differently. For example, if someone's name is "Smith" but is misspelled as "Smyth," a phonetic function would convert both spellings into a code that represents how they sound. This way, the two names can still be matched, even though the spellings are different. Common phonetic algorithm that is used for English is **Soundex**.

By setting `phonetic` and specifying the phonetic algorithm to `phonfun`, phonetic algorithm will be applied while doing deduplication:

```{r}
rpairs<- compare.dedup(RLdata500, phonetic = 1, phonfun = soundex, blockfld = 1)
rpairs$pairs[1:5, ]
```

String comparators are another approach that measures how similar two strings are. Instead of converting the words into phonetic codes, string comparators compare the strings directly and give a similarity score between 0 and 1. A score of 0 means the strings are completely different, and a score of 1 means they are exactly the same. Commonly used string comparison functions are **jarowinkler** (Winkler (1990)) and edit distance-based **levenshteinSim**(Levenshtein).

String comparators will be applied by setting the `strcmp` argument to `TRUE` and specifying a comparison function using the `strcmpfun` argument:

```{r}
rpairs <- compare.dedup(RLdata500, strcmp = TRUE, strcmpfun = jarowinkler, blockfld = 1)
rpairs$pairs[1:5, ]
```
